### 2020-10-11 shiny/HeatEqn.Rmd

---
title: "Heat Equation"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eruptions, echo=FALSE}
library(shiny)
library(ggplot2)

# PDE
Laplacian <- function(y_k, dx){
  delta <- NULL
  len = length(y_k)
  # boundary condition
  delta[1] =(y_k[2]-y_k[1])/dx^2 
  delta[len] = (y_k[len-1]- y_k[len])/dx^2
  delta[2: (len-1)] <- diff(diff(y_k)/dx) /dx
  return(delta)
}
HeatEqn <- function(y0, dx, C, sd, dt, Nt, history = TRUE){
  yhat = list()
  yhat[[1]] = y0
  for (i in 1:(Nt+1)) {
    g = C* yhat[[i]]/sd
    yhat[[i+1]] = yhat[[i]] + g* Laplacian(yhat[[i]], dx) * dt
  }
  ifelse(history, return(yhat), return(yhat[[Nt+2]]))
}

###################################################################

# Shinny

ui <- fluidPage(

  # Application title
  #titlePanel("Data Generator"),

  # Sidebar with a slider input for number of bins 
  fluidRow(
    column(4, 
      textInput(inputId = "seed",
                label = "Choose set.seed",
                value = "1"),

      textInput(inputId = "equation", 
                label = "Underlying function", 
                value = "x^2"),

      sliderInput("xrange",
                  "Range of x:",
                  min = -10,
                  max = 10,
                  value = c(-1,1) )
    ),
    column(4,
      sliderInput(inputId = "nobs",
                  label = "Number of observations",
                  min = 2,
                  max = 300,
                  value = 20),

      sliderInput(inputId = "mean",
                  label = "Mean of noise",
                  min = -10,
                  max = 10,
                  value = 0.25,
                  step = 0.01),

      textInput(inputId = "sdtext", 
                label = "Standard deviation of noise",
                value = "0.1")
     ),
    column(4, 
      sliderInput(inputId = "dt",
                  label = "Set value for dt",
                  min = 0,
                  max = 0.0001,
                  value = 0.0001),

      sliderInput(inputId = "Nt",
                  label = "Set value for Nt",
                  min = 1,
                  max = 4000,
                  value = 200),
      #textInput(inputId = "g", 
      #          label = "Value of g",
      #          value = "x^2")
      sliderInput(inputId = "C",
                  label = "Set value for C",
                  min = 0.01,
                  max = 10,
                  value = 1)
     )
  ),

  fluidRow(
    column(6, 
           plotOutput("Plot1")),
    column(6, 
           plotOutput("Plot2")
           )
  )
)

###################################################################

server <- function(input, output){

  equation = function(x){ eval(parse(text = input$equation)) }

   sdtext = function(x){ eval(parse(text = input$sdtext)) }

  # diffusi = function(x){ eval(parse(text = input$g)) }
  C =  function(x){ eval(parse(text = input$C)) }

  xvalue = reactive({ seq(input$xrange[1], input$xrange[2], length.out = input$nobs) })

  noise = reactive({
    set.seed(input$seed)
    rnorm(input$nobs, mean = input$mean, sd = input$sdtext)})

  noise2 = reactive({
    set.seed(input$seed)
    rnorm(input$nobs, mean = input$mean, sd = sdtext(xvalue()))})

  yvalue = reactive({ equation(xvalue()) + noise2() })

  #  hqvalue = reactive({
  #     HeatEqn(yvalue(), 0.001, g = input$g, dt = input$dt, Nt= input$Nt, history = F)
  #})

  output$Plot1 <- renderPlot({
    x = xvalue()
    dx = x[2] - x[1]
    y = yvalue()
    y1 = equation(x)
    hq = HeatEqn(y, dx, input$C, noise2()+0.001, input$dt, input$Nt, history = F)

    df <- data.frame(cbind(x,y,y1,hq))
    ggplot(df) + 
      geom_point(aes(y = y, x = x, color = '1')) +
      geom_line(aes(y = y1, x = x, color = '2'), linetype = "twodash") + 
      geom_line(aes(y = hq, x = x, color = '3')) + 
      scale_color_discrete(name = "Y", labels = c("observations", "real value", "fitted value")) +
      ggtitle("One Dimensional Finite Difference Method")

  })

  output$Plot2 <- renderPlot({
    x = xvalue()
    dx = x[2] - x[1]
    y = yvalue()
    y1 = equation(x)
    hq = HeatEqn(y, dx, input$C, noise2()+0.001, input$dt, input$Nt, history = F)
    ratio = y1 /(noise2()+0.001)
    g = ratio * input$C
    df2 <- data.frame(cbind(x,ratio,g))
    ggplot(df2) + 
      geom_line(aes(y = ratio, x = x, color = '1'), linetype = "twodash") + 
      geom_line(aes(y = g, x = x, color = '2')) + 
      scale_color_discrete(name = "Y", labels = c("ratio", "g")) +
      ggtitle("Smoothing Strategy: Tuning of g")

  })

}

shinyApp(ui = ui, server = server)
```